Safe Transferable Regions

### 2016-10-14 paper/abstract.tex
various

Manual memory management is susceptible to the usual perils of
dangling pointers. A key contribution of this paper is a
refinement-based region type system, for a C\#-like typed object-oriented
language with higher-order functions and generics, that ensures the
memory safety of programs in the presence of transferable regions.
We complement our type system with a type inference algorithm,
which eliminates the need for programmers to write region annotations on types.
The cornerstone of our inference algorithm is a novel constraint solver that
performs abduction in a partial-order constraint domain to infer weakest solutions
to recursive constraints.
We establish soundness and partial completeness results for the type inference algorithm:
in particular, the key steps of constraint-generation and constraint-solving
in the type inference algorithm are complete.

>>>>

Manual memory management is susceptible to the usual perils of
dangling pointers. A key contribution of this paper is a
refinement-based region type system, for a C\#-like typed object-oriented
language with higher-order functions and generics, that ensures the
memory safety of programs in the presence of transferable regions.
We complement our type system with a type inference algorithm, which eliminates the need for programmers to write region annotations on types.
% The cornerstone of our inference algorithm is a novel constraint solver that
% performs abduction in a partial-order constraint domain to infer weakest solutions
% to recursive constraints.
We establish soundness and partial completeness results for the type inference algorithm: in particular, the key steps of constraint-generation and constraint-solving in the type inference algorithm are complete. We describe a prototype implementation of our inference algorithm and the constraint solver, and our experience of using it to enforce memory safety in dataflow programs.

### 2016-10-14 paper/type-inference.tex
proofs and space compression
various

\newcommand{\saturate}[1]{{#1}^*}
\newcommand{\myground}[1]{\saturate{#1}_g}
\newcommand{\consOf}[1]{\textsc{GenConstraint}(#1)}
\newcommand{\satC}{\saturate{C}}
\newcommand{\groundC}{\myground{C}}
\newcommand{\rhoC}{C_\rho}
\newcommand{\solveCon}[1]{\textsc{Solve}(#1)}

@@

@@

Every predicate variable $\varphi$ denotes an unknown \emph{region-constraint},
%% over a set of fixed formal region parameters,
where the set of region-constraints $\phi$ is defined by:
\begin{smathpar}
\begin{array}{lcl}
\phi & \coloneqq & true \ALT \rho \outlives \rho \ALT \phi \conj \phi \\
\end{array}
\end{smathpar}
We will use $\rho_1 = \rho_2$ as shorthand for $\rho_1 \outlives \rho_2 \conj \rho_2 \outlives \rho_1$.
We will also use a set $S$ of primitive constraints of the form $\rho_1 \outlives \rho_2$
as shorthand for $\conj_{\phi \in S} \phi$.

We will use the term \emph{validity constraint} to denote an entailment constraint
of the form $\isvalid{\varphi}{\rho_1 \outlives \rho_2}$.

Our constraints also make uses of \emph{pending substitutions} $F$:
% \footnote{we borrowed this terminology from~\cite{ltpldi08}}
A pending substitution serves to bind formal region parameters in $\varphi$ to the actual region parameters
used in a particular context:
\begin{smathpar}
\begin{array}{lcl}
F & \coloneqq & \cdot \ALT [\rho/\rho]F \\
\end{array}
\end{smathpar}
E.g., in the validity constraint $\isvalid{\rgn_1 \outlives
\rgn_2}{[\rgn_1/\rho_1][\rgn_2/\rho_2]\varphi}$, the pending substitution
is $[\rgn_1/\rho_1][\rgn_2/\rho_2]$. Any concrete formula (call it
$\phisol$) over variables $\rho_1$ and $\rho_2$ is a solution to
$\varphi$ if and only if the formula obtained by substituting $\rgn_1$
and $\rgn_2$ for $\rho_1$ and $\rho_2$ (resp.) in $\phisol$ is
deducible from $\rgn_1 \outlives \rgn_2$.

>>>>

\newcommand{\saturate}[1]{{#1}^*}
\newcommand{\myground}[1]{\saturate{#1}_g}
\newcommand{\consOf}[1]{\textsc{GenConstraint}(#1)}
\newcommand{\satC}{\saturate{C}}
\newcommand{\groundC}{\myground{C}}
\newcommand{\rhoC}{\text{WF}_R}
\newcommand{\solveCon}[1]{\textsc{Solve}(#1)}

@@

\input{constraint-syntax}

@@

Every predicate variable $\varphi$ denotes an unknown \emph{region-constraint},
over a set of fixed formal region parameters.
%% where the set of region-constraints $\phi$ is defined by:
%% \begin{smathpar}
%% \begin{array}{lcl}
%% \phi & \coloneqq & true \ALT \rho \outlives \rho \ALT \phi \conj \phi \\
%% \end{array}
%% \end{smathpar}
%% We will use $\rho_1 = \rho_2$ as shorthand for $\rho_1 \outlives \rho_2 \conj \rho_2 \outlives \rho_1$.
%% We will also use a set $S$ of primitive constraints of the form $\rho_1 \outlives \rho_2$
%% as shorthand for $\conj_{\phi \in S} \phi$.
%% 
We will use the term \emph{validity constraint} to denote an entailment constraint
of the form $\isvalid{\varphi}{\rho_1 \outlives \rho_2}$.

Our constraints also make uses of \emph{pending substitutions} $F$:
A pending substitution serves to bind formal region parameters in $\varphi$ to the actual region parameters
used in a particular context:
%% \begin{smathpar}
%% \begin{array}{lcl}
%% F & \coloneqq & \cdot \ALT [\rho/\rho]F \\
%% \end{array}
%% \end{smathpar}
E.g., in the validity constraint $\isvalid{\rgn_1 \outlives
\rgn_2}{[\rgn_1/\rho_1][\rgn_2/\rho_2]\varphi}$, the pending substitution
is $[\rgn_1/\rho_1][\rgn_2/\rho_2]$. Any concrete formula (call it
$\phisol$) over variables $\rho_1$ and $\rho_2$ is a solution to
$\varphi$ if and only if the formula obtained by substituting $\rgn_1$
and $\rgn_2$ for $\rho_1$ and $\rho_2$ (resp.) in $\phisol$ is
deducible from $\rgn_1 \outlives \rgn_2$.
